---
title: "DataExplorer"
author: "Sally Heo"
date: "4/10/2022"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: yeti
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(fig.path = "../output/", echo = TRUE)
```

## Load libraries and data ##
```{r, warning=FALSE, message=FALSE}
library(tidyverse) #Load tidyverse
library(here) #Load here
library(DataExplorer) #Load DataExplorer
library(ggthemes) #Load ggthemes
```
## Load Data ##
```{r, warning=FALSE, message=FALSE}
chem <- read_csv(here("DataExplorer", "data", "chemicaldata_maunalua.csv"))
View(chem)
```
## Function: plot_correlation ##
* This function creates a correlation heatmap for discrete features 
* Discrete data is quantitative data that relies on counts in whole numbers or decimals.
* Analyze correlation in your data

```{r, warning=FALSE, message=FALSE}
plot_correlation(chem,#Input the data, in this case it is chem
                 type = c("all", "discrete", "continuous"), #The column types to be included in the correlation plot; "All" = all columns, "discrete" is for discrete features, "continuous" is for continuous. 
                 maxcat = 20L, #Maximum categories allowed for each discrete feature
                 cor_args = list(),# List of arguments to be passed to the plot correlation function
                 geom_text_args = list(), #List of arguments to geom_text
                 title = NULL, #Plot title, NULL is the default (no title)
                 ggtheme = theme_classic(), #Setting theme_classic
                 theme_config = list(legend.position = "bottom", axis.text.x = element_text(angle = 90))#Adjusting axis text, and legend position
                 )
```

## Function(a): create_report() ##
* This function creates a data profiling report for our Chem Data
* Organizes the data profiling statistics into a report
* This is a really important function and runs most of the explatory data analysis (EDA) functions and gives a output html file. 

## Function (b): configure_report()
* This function allows us to customize each individual section of the data profiling report
* We can turn off sections of the data report that we don't want to show
* We can also set a global theme
* For example for add_plot qq = FALSE, if we change FALSE to TRUE, the qq plot will be included in the data profiling report.
* We can also configure a report where only QQ and box plots are included and the correlation analysis is not added.

## Arguments ##
* add_plot_qq: Specifies whether or not the QQ plot is added in the report
* add_plot_boxplot: Specifies whether or not the boxplot is added in the report
* add_plot_histogram: Specifies whether or not the histogram is added to the report
* add_plot_correlation: Specifies whether or not the correlation analysis is added to the report
* global_ggtheme: The global setting and to set the theme of the report (ie theme_minimal, theme_gray, theme_few) 

```{r}
config <- configure_report( #config argument, the report configuration generated by configure_report
  add_plot_qq = TRUE, #Show qq plot in the data profiling report
  add_plot_boxplot = TRUE, #show the boxplot in the data profiling report
  add_plot_histogram = TRUE, #show the histogram in the data profiling report
  add_plot_correlation = TRUE, #show the correlation analysis in the data profiling report
  global_ggtheme = quote(theme_gray(base_size = 14)) #Setting global setting for the theme of the report
)
create_report(chem, config = config) #creates a data profiling report
```
```{r}


```





